Geometric variational problems

### 2020-11-15 NSF2020/Description/newspecificaim3.tex
fixed typos in SA 2 and 3

\subsubsection{Vesicles in background flows}
Numerically solving the Stokes and screened Laplace system~\eqref{SL}
and~\eqref{eq:stokes} is nontrivial, and we will develop efficient,
robust, high-order in time and space algorithms suited to the problem.
The challenges are: {\bf (1)} The equations express a two-way coupling
since the flow changes the position of the suspended particles,
and in return the geometry-dependent hydrophobic stresses impart a force
on the flow. {\bf (2)} The inputs to the Stokes equations are particle
configurations, forces, and torques. The outputs are the rigid body
translation and angular velocities used to update the particle positions
(see Figure~\ref{fig:flow_map}). Although the underlying equations are
linear, the overall problem is highly nonlinear because the domain is
constantly changing. {\bf (3)} Self-assembly causes the particles to
come into close contact. As a result, an exceptional spatial accuracy is
required to resolve the fields between adjacent particles. Finally, the
physically relevant elastic properties of bilayer become apparent at
large length scales and for large particle-numbers, which increases the
computational complexity of our simulations.

We include a background flow by replacing the third equation in
\eqref{eq:stokes} with the condition $\mathbf{u}(\mathbf{x})
\to \mathbf{u}_{\infty}(\mathbf{x})$ as $|\mathbf{x}| \to
\infty$. To incorporate the far-field flow, we use the representation 
\begin{align}
\label{PowerMiranda}
  {\bf u} = {\bf u}_{\infty} + K\boldsymbol{\eta} + 
    \sum_{i=1}^N \left(S(\cdot, {\bf a}_i) {\bf F}_i + 
                 R(\cdot, {\bf a}_i) {\bf G}_i\right),
\end{align}
where $S$ and $R$ are stokeslets and rotlets supported at the respective
particle centers~\cite{leal_2007} and $K\boldsymbol{\eta}$ is a layer
potential for the unknown density function $\boldsymbol{\eta}$. With the
exception of the rigid body conditions, the
representation~\eqref{PowerMiranda} automatically satisfies all the
hydrodynamic equations including the far-field condition. The rigid
motion conditions follow by requiring the viscous stress vanishes across
the particle boundaries.
%
We will compare the behavior of our particle-based vesicles in Stokes
flow to well-established models that enforce area incompressibility and
volume conservation in vesicle membrane
dynamics~\cite{torres-sanchez_millan_arroyo_2019,mahapatra_saintillan_rangamani_2020, Steigmann99, C6SM02452A}.
Some preliminary results are in \S \ref{sec:preliminary_work}. 
%
%Our
%preliminary results show that the particle-based bilayers have the same
%large area modulus as real lipid bilayer membranes, so we expect
%realistic results from our flow simulations. Under moderate shear rates,
%the vesicle elongates and takes on the tank-treading elliptical shape.
%We can directly check for area compressibility and volume conservation
%properties from particle simulations by tracking the change in area
%density of the midplane as a function of shear flow rate, along with
%evaluating the tangential divergence of the velocity.
%
%Bilayer membranes have a small permeability to
%water~\cite{323e9a2f0c58487ea82518d7a1f96485}, and modelers often assume
%a vesicle membrane that conserves volume. In our particle-based
%approach, water motion across the bilayer is limited by the size of the
%gaps between the particles, which is an artifact of coarse-graining.
%Making these gaps smaller comes at the expense of numerical accuracy,
%and we will assess if there is a reasonable trade-off between
%approximate volume conservation and simulation complexity. 
%
%Finally, the particle-based vesicle bilayers have two distinct leaflets
%consisting of the particles in contact with the vesicle interior, and
%those in contact with the surrounding fluid. Since there is a nonzero
%separation between these layers, an important effect we observe is that
%the leaflets slide against one another under shear flow. This implies
%that part of the viscous dissipation in the aqueous phase is enhanced by
%intermonolayer friction~\cite{SHKULIPA2005823, ShkulipaThesis}.
%Intermonolayer slip enters zero-thickness membrane models as a velocity
%jump boundary condition and leads to multiple solution
%states~\cite{schwalbe_vlahovska_miksis_2010}. In the present proposal,
%intermonolayer slip is an artifact of monolayer independence and we can
%control for slip as a function of shear rate, vesicle diameter, and
%particle geometry.

>>>>

\subsubsection{Vesicles in background flows}
Numerically solving the Stokes and screened Laplace system~\eqref{SL}
and~\eqref{eq:stokes} is nontrivial, and we will develop efficient,
robust, high-order-in-time and -space algorithms suited to the problem.
There are a few challenges. {\bf (1)} The equations express a two-way coupling
since the flow changes the position of the suspended particles,
and in return, the geometry-dependent hydrophobic stresses impart a force
on the flow. {\bf (2)} The inputs to the Stokes equations are particle
configurations, forces, and torques. The outputs are the rigid body
translation and angular velocities used to update the particle positions
(see Figure~\ref{fig:flow_map}). Although the underlying equations are
linear, the overall problem is highly nonlinear because the domain is
constantly changing. {\bf (3)} Self-assembly causes the particles to
come into close contact. As a result, an exceptional spatial accuracy is
required to resolve the fields between adjacent particles. Furthermore, 
physically relevant elastic properties of bilayers become apparent at
large length scales and for large particle-numbers, which increases the
computational complexity of our simulations.

We include a background flow by replacing the third equation in
\eqref{eq:stokes} with the condition $\mathbf{u}(\mathbf{x})
\to \mathbf{u}_{\infty}(\mathbf{x})$ as $|\mathbf{x}| \to
\infty$. To incorporate the far-field flow, we use the representation 
\begin{align}
\label{PowerMiranda}
  {\bf u} = {\bf u}_{\infty} + L\boldsymbol{\eta} + 
    \sum_{i=1}^N \left(S(\cdot, {\bf a}_i) {\bf F}_i + 
                 R(\cdot, {\bf a}_i) {\bf G}_i\right),
\end{align}
where $S$ and $R$ are stokeslets and rotlets supported at the respective
particle centers~\cite{leal_2007} and $L\boldsymbol{\eta}$ is a layer
potential for the unknown density function $\boldsymbol{\eta}$. With the
exception of the rigid body conditions, the
representation~\eqref{PowerMiranda} automatically satisfies all the
hydrodynamic equations including the far-field condition. The rigid
body conditions follow by requiring the viscous stresses to vanish across
the particle boundaries.
%
We will compare the behavior of our particle-based vesicles in Stokes
flow to well-established models that enforce area incompressibility and
volume conservation in vesicle membrane
dynamics~\cite{torres-sanchez_millan_arroyo_2019,mahapatra_saintillan_rangamani_2020, Steigmann99, C6SM02452A}.
\S \ref{sec:preliminary_work} contains our preliminary results.
%
%Our
%preliminary results show that the particle-based bilayers have the same
%large area modulus as real lipid bilayer membranes, so we expect
%realistic results from our flow simulations. Under moderate shear rates,
%the vesicle elongates and takes on the tank-treading elliptical shape.
%We can directly check for area compressibility and volume conservation
%properties from particle simulations by tracking the change in area
%density of the midplane as a function of shear flow rate, along with
%evaluating the tangential divergence of the velocity.
%
%Bilayer membranes have a small permeability to
%water~\cite{323e9a2f0c58487ea82518d7a1f96485}, and modelers often assume
%a vesicle membrane that conserves volume. In our particle-based
%approach, water motion across the bilayer is limited by the size of the
%gaps between the particles, which is an artifact of coarse-graining.
%Making these gaps smaller comes at the expense of numerical accuracy,
%and we will assess if there is a reasonable trade-off between
%approximate volume conservation and simulation complexity. 
%
%Finally, the particle-based vesicle bilayers have two distinct leaflets
%consisting of the particles in contact with the vesicle interior, and
%those in contact with the surrounding fluid. Since there is a nonzero
%separation between these layers, an important effect we observe is that
%the leaflets slide against one another under shear flow. This implies
%that part of the viscous dissipation in the aqueous phase is enhanced by
%intermonolayer friction~\cite{SHKULIPA2005823, ShkulipaThesis}.
%Intermonolayer slip enters zero-thickness membrane models as a velocity
%jump boundary condition and leads to multiple solution
%states~\cite{schwalbe_vlahovska_miksis_2010}. In the present proposal,
%intermonolayer slip is an artifact of monolayer independence and we can
%control for slip as a function of shear rate, vesicle diameter, and
%particle geometry.

### 2020-11-15 NSF2020/Description/preliminarywork.tex
fixed typos in SA 2 and 3

\begin{wrapfigure}[11]{l}{0.22\textwidth}
  \centering{\includegraphics[width=0.22\textwidth]{Figures/PW_fig5.pdf}}
  \caption{\label{fig:flow_rupture} \footnotesize Rupture of a
  two-dimensional vesicle at large shear rates.}
\end{wrapfigure}
The above considerations further justify using the particle-based model
to study material properties as outlined in
\S\ref{subsec:specific_aim_1}. Furthermore, the particle system allows
for intermonolayer slip~\cite{SHKULIPA2005823, ShkulipaThesis}. In
Figure \ref{fig:tank_treading}B, the inner angular velocity is slightly
less than the outer angular velocity. Intermonolayer slip enters
zero-thickness membrane models as a velocity jump boundary condition
\cite{schwalbe_vlahovska_miksis_2010}, for example, but in the present
model, it is a consequence of monolayer independence. For large shear
rates, viscous forces exceed the hydrophobic attraction and the vesicle
ruptures; the flow carries away segmented membrane patches (Figure
\ref{fig:flow_rupture}).

>>>>

\begin{wrapfigure}[11]{l}{0.22\textwidth}
  \centering{\includegraphics[width=0.22\textwidth]{Figures/PW_fig5.pdf}}
  \caption{\label{fig:flow_rupture} \footnotesize Rupture of a
  two-dimensional vesicle at large shear rates.}
\end{wrapfigure}
The above considerations further justify using the particle-based model
to study material properties as outlined in
\S\ref{subsec:specific_aim_1}. Furthermore, the particle system allows
for intermonolayer slip~\cite{SHKULIPA2005823, ShkulipaThesis}. In
Figure \ref{fig:tank_treading}B, the inner angular velocity is slightly
less than the outer angular velocity. Intermonolayer slip enters
zero-thickness membrane models as a velocity jump boundary condition
\cite{schwalbe_vlahovska_miksis_2010}, for example, but in the present
model, it is a consequence of monolayer independence. For large shear
rates, viscous forces exceed the hydrophobic attraction to the point that
the vesicle ruptures. The flow carries away segmented membrane patches
(Figure \ref{fig:flow_rupture}).
